Model-based variable selection for regression with highly correlated variables.

### 2014-11-03 www/article.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@902 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\paragraph{Remarks}
\begin{itemize}
\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
\item As a consequence of Hypothesis~\ref{H1} and~\ref{H2} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals. Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}. 
\end{itemize}

@@

In practice, the strategy we propose is to rely estimate of  $\hat{\boldsymbol{\beta}}$ upon Equation~(\ref{eq:modexpl}). The practitioner can choose any estimate of its choice, like {\sc ols} or any variable selection procedure like {\sc lasso}. In other words, it is possible to see $(\ref{eq:modexpl})$ as a kind of {\it pretreatment} for decorrelating covariates, while assuming nothing on the subsequent estimate process.

In the following, we will denote by {\sc CorReg} the new proposed strategy.

@@

\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_{J_r}^*=\boldsymbol{\beta}_{r}^*={\bf 0}$ and $\boldsymbol{\beta}_{J_f}^*=\boldsymbol{\beta}_{f}^*$, (\ref{eq:modexpl}) can be rewritten
\begin{equation}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
\end{equation}
The {\sc ols} estimate of $\boldsymbol{\beta}^*$ is then given by
	\begin{equation}
		\hat{\boldsymbol{\beta}}^*_f=\left(\boldsymbol{X}_f'\boldsymbol{X}_f \right) ^{-1}\boldsymbol{X}_f'\boldsymbol{Y} \quad \mbox{and} \quad \hat{\boldsymbol{\beta}}^*_r = {\bf 0}.
	\end{equation}
As usual, the {\sc ols} estimate $\hat{\boldsymbol{\beta}}^*$ of $\boldsymbol{\beta}^*$ is unbiased but, however, contrary to $\hat{\boldsymbol{\beta}}$, it could be a biased estimate of $\boldsymbol{\beta}$ since
		\begin{equation}
			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^* \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
		\end{equation}
In return, its variance could be reduced compared to this one of $\hat{\boldsymbol{\beta}}$ given in (\ref{eq:varOLS}) as soon as values of $\sigma_j$ are small enough (it means strong correlations in sub-regressions) as we can see in the following expression
		\begin{equation}
			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}. \label{eq:varOLS*}
		\end{equation}
Indeed, no correlations between covariates $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditioned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}). This bias-variance trade off can be resumed by the Mean Squared Error (\textsc{mse}) associated to both estimates:
	\begin{eqnarray}
			\textsc{mse}(\hat{\boldsymbol{\beta}})&=& \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}),
			 \\
			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j =1}^{d_r}\sigma_{j}\beta_{J_r^j} \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
	\end{eqnarray}

@@

\begin{hyp}\label{H4}
All variates $\boldsymbol{X}^j$ with $j \in J_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
\begin{equation}
\mathbb{P}(\boldsymbol{X}_f^j|\boldsymbol{S};\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j) = \sum_{h=1}^{k_j} \pi_{hj} \mathcal{N}_n(\mu_{hj} . (1,\ldots,1)',\lambda_{hj}^2\boldsymbol{I}),
\end{equation}
where $\boldsymbol{\pi}_{j}=(\pi_{1j},\ldots,\pi_{k_jj})$ is the vector of mixing proportions with $\forall 1\leq h\leq k_j , \pi_{hj}>0$ and $\sum_{h=1}^{k_j}\pi_{hj}=1$, $\boldsymbol{\mu}_j=(\mu_{1j},\ldots,\mu_{k_jj})$  is the vector of centers and $\boldsymbol{\lambda}^2_j=(\lambda^2_{1j},\ldots,\lambda^2_{k_jj})$ is the vector of variances. We stack together all the mixture parameters in $\boldsymbol{\theta}=(\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j ; j \in J_f)$.
\end{hyp}

@@

We have generated a non-linear sub-regression to test the robustness of our model. $\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as before. We design a unique, possibly non linear, sub-regression $\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2+\boldsymbol{X}_3+ \varepsilon_1$. The matrix $\boldsymbol{X}$ is then scaled and we set $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\varepsilon_Y$. We let $a$ vary between $0$ and $10$ to increase progressively the non-linear part of the sub-regression.
Figure~\ref{resnonlin} shows that the {\sc mcmc} algorithm has more difficulties to find a linear structure (by  $\mbox{{\sc bic}}_H$ as the non-linear part of the sub-regression increases with $a$. But the model is quite robust, preserving good efficient for small values of $a$.
In addition, Figure~\ref{MSEnonlin} illustrates the advantage of using {\sc CorReg}, even with non-linear sub-regressions, concerning the quality of the {\sc mse}.

>>>>

\paragraph{Remarks}
\begin{itemize}
\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
%\item As a consequence of Hypothesis~\ref{H1} and~\ref{H2} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated.
\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals. Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}. 
\end{itemize}

@@

In practice, the strategy we propose is to rely estimate of  $\hat{\boldsymbol{\beta}}$ upon Equation~(\ref{eq:modexpl}). The practitioner can choose any estimate of its choice, like {\sc ols} or any variable selection procedure like {\sc lasso}. In other words, it is possible to see $(\ref{eq:modexpl})$ as a kind of {\it pretreatment} for decorrelating covariates, while assuming nothing on the subsequent estimate process.
\paragraph{Remark}
\begin{itemize}
%\item Sub-regressions defined in (\ref{eq:SR}) are very easy to understand by any practitioner and, thus, will give a clear view of all the correlations present in the dataset at hand.
\item As a consequence of Hypothesis~\ref{H1} to~\ref{H3} , ``free'' covariates $\boldsymbol{X}_f$ are {\it all} decorrelated (see the Lemma in the Appendix about identifiability).
%\item We have considered correlations between the covariates of the main regression on $\boldsymbol{Y}$, not between the residuals. Thus $\boldsymbol{S}$ does not depend on $\boldsymbol{Y}$ and it can be estimated independently as we will see in Section~\ref{sec:estimateS}.
%\item The model of sub-regressions $\boldsymbol{S}$ gives a system of linear regressions that can be viewed as a recursive Simultaneous Equation Model (\textsc{sem})\cite{davidson1993estimation,TIMM} or also as a Seemingly Unrelated Regression (\textsc{sur}) \cite{SURzellner}. 
\end{itemize} 
In the following, we will denote by {\sc CorReg} the new proposed strategy.

@@

\paragraph{Qualitative illustration} Noting $\boldsymbol{\beta}^*$ the vector of coefficients of the same dimension as $\boldsymbol{\beta}$ with $\boldsymbol{\beta}_{J_r}^*=\boldsymbol{\beta}_{r}^*={\bf 0}$ and $\boldsymbol{\beta}_{J_f}^*=\boldsymbol{\beta}_{f}^*$, (\ref{eq:modexpl}) can be rewritten
\begin{equation}
	\boldsymbol{Y}{|\boldsymbol{X}_f,\boldsymbol{S}};\boldsymbol{\beta},\boldsymbol{\alpha},\sigma_Y^2,\boldsymbol{\sigma}^2=\boldsymbol{X}_f\boldsymbol{\beta}_f^*+\boldsymbol{X}_r\boldsymbol{\beta}_r^*+\boldsymbol{\varepsilon}_Y^*.\label{eq:modexpl2}
\end{equation}
The {\sc ols} estimate of $\boldsymbol{\beta}^*$ is then given by
	\begin{equation}
		\hat{\boldsymbol{\beta}}^*_f=\left(\boldsymbol{X}_f'\boldsymbol{X}_f \right) ^{-1}\boldsymbol{X}_f'\boldsymbol{Y} \quad \mbox{and} \quad \hat{\boldsymbol{\beta}}^*_r = {\bf 0}.
	\end{equation}
As usual, the {\sc ols} estimate $\hat{\boldsymbol{\beta}}^*$ of $\boldsymbol{\beta}^*$ is unbiased but, however, contrary to $\hat{\boldsymbol{\beta}}$, it could be a biased estimate of $\boldsymbol{\beta}$ since
		\begin{equation}
			\mathbb{E}(\hat{\boldsymbol{\beta}}_f^*)=\boldsymbol{\beta}_f+\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^* \quad \textrm{and}\quad \mathbb{E}(\hat{\boldsymbol{\beta}}_r^*)=\boldsymbol{0}.
		\end{equation}
In return, its variance could be reduced compared to this one of $\hat{\boldsymbol{\beta}}$ given in (\ref{eq:varOLS}) as soon as values of $\sigma_j$ are small enough (it means strong correlations in sub-regressions) as we can see in the following expression
		\begin{equation}
			\operatorname{Var}(\hat{\boldsymbol{\beta}}_f^*)= (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2)(\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1} \quad \textrm{and} \quad\operatorname{Var}(\hat{\boldsymbol{\beta}}_r^*)= \boldsymbol{0}. \label{eq:varOLS*}
		\end{equation}
Indeed, no correlations between covariates $\boldsymbol{X}_f$ imply that the matrix $\boldsymbol{X}_f' \boldsymbol{X}_f$ could be sufficiently better conditioned than the matrix $\boldsymbol{X}' \boldsymbol{X}$ involved in (\ref{eq:varOLS}) to balance the added variance $\sum_{j \in I_r}\sigma^2_{j}\beta_{j}^2$ in (\ref{eq:varOLS*}). This bias-variance trade off can be resumed by the Mean Squared Error (\textsc{mse}) associated to both estimates:
	\begin{eqnarray}
			\textsc{mse}(\hat{\boldsymbol{\beta}})&=& \sigma_Y^2 \operatorname{Tr}((\boldsymbol{X}'\boldsymbol{X})^{-1}),
			 \\
			\textsc{mse}(\hat{\boldsymbol{\beta}}^*)&=& \parallel\sum_{j =1}^{d_r}\beta_{J_r^j}\boldsymbol{\alpha}_j^* \parallel_2^2 +\parallel \boldsymbol{\beta}_r\parallel^2_2 + (\sigma^2_Y+\sum_{j =1}^{d_r}\sigma^2_{j}\beta_{J_r^j}^2 ) \operatorname{Tr}((\boldsymbol{X}_f' \boldsymbol{X}_f)^{-1}).
	\end{eqnarray}

@@

\begin{hyp}\label{H4}
All covariates $\boldsymbol{X}^j$ with $j \in J_f$ are mutually independent and arise from the following Gaussian mixture of $k_j$ components
\begin{equation}
\mathbb{P}(\boldsymbol{X}_f^j|\boldsymbol{S};\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j) = \sum_{h=1}^{k_j} \pi_{hj} \mathcal{N}_n(\mu_{hj} . (1,\ldots,1)',\lambda_{hj}^2\boldsymbol{I}),
\end{equation}
where $\boldsymbol{\pi}_{j}=(\pi_{1j},\ldots,\pi_{k_jj})$ is the vector of mixing proportions with $\forall 1\leq h\leq k_j , \pi_{hj}>0$ and $\sum_{h=1}^{k_j}\pi_{hj}=1$, $\boldsymbol{\mu}_j=(\mu_{1j},\ldots,\mu_{k_jj})$  is the vector of centers and $\boldsymbol{\lambda}^2_j=(\lambda^2_{1j},\ldots,\lambda^2_{k_jj})$ is the vector of variances. We stack together all the mixture parameters in $\boldsymbol{\theta}=(\boldsymbol{\pi}_{j},\boldsymbol{\mu}_j,\boldsymbol{\lambda}^2_j ; j \in J_f)$.
\end{hyp}

@@

We have generated a non-linear sub-regression to test the robustness of our model. $\boldsymbol{X}_f$ is a set of 6 independent Gaussian mixtures defined as before. We design a unique, possibly non linear, sub-regression $\boldsymbol{X}_7=a\boldsymbol{X}_1^2+\boldsymbol{X}_2+\boldsymbol{X}_3+ \boldsymbol{\varepsilon}_1$. The matrix $\boldsymbol{X}$ is then scaled and we set $\boldsymbol{Y}=\sum_{i=1}^7\boldsymbol{X}_i+\boldsymbol{\varepsilon}_Y$. We let $a$ vary between $0$ and $10$ to increase progressively the non-linear part of the sub-regression.
Figure~\ref{resnonlin} shows that the {\sc mcmc} algorithm has more difficulties to find a linear structure (by  $\mbox{{\sc bic}}_H$ as the non-linear part of the sub-regression increases with $a$. But the model is quite robust, preserving good efficient for small values of $a$.
In addition, Figure~\ref{MSEnonlin} illustrates the advantage of using {\sc CorReg}, even with non-linear sub-regressions, concerning the quality of the {\sc mse}.
